---
title: "Amino Acid"
author: "Richard Edwards"
date: "June 4, 2015"
output: html_document
---


Coding for the Side Chain group of the 20 common Amino Acids
```{r}
#Amino Acids                                         NH2
 #                                                   |
 #All Amino Acids have the general composition of   HC--C(O)OH with the 'R' group
 #                                                   |         specifying the 
 #                                                   R         amino acid residue
 
 #Defining Aromatic 'R' Groups
 Phenyl <- c("C",rep("CH",5))
 Imidazole <- c("C","CH","NH","CH","N+H")
 Indole <- c("C","CH","NH","C","C",rep("CH",4))
 Phenol <- c("C","CH","CH","C_OH","CH","CH")
 #The following will relate the 1-letter abreviation of 
 #the 20 AA residues with their respecitve 'R' groups

A <- Ala <- c("CH3")                              #Alanine (Ala)
C <- Cys <- c("CH2", "SH")                        #Cysteine (Cys)
D <- Asp <- c("CH2","COO-")                       #Aspartate (Asp)
E <- Glu <- c("CH2","CH2","COO-")                 #Glutamate (Glu)
F <- Phe <- c("CH2",Phenyl)                       #Phenylalanine (Phe)
G <- Gly <- c("H")                                #Glycine (Gly)
H <- His <- c("CH2",Imidazole)                    #Histidine (His)
I <- Ile <- c("Methyl_CH","CH2","CH3")            #IsoLeucine (Ile)
K <- Lys <- c(rep("CH2",4), "N+H3")               #Lysine (Lys)
L <- Leu <- c("CH2","Methyl_CH","CH3")            #Leucine (Leu)
M <- Met <- c("CH2","CH2","S","CH3")              #Methionine (Met)
N <- Asn <- c("CH2","CO","NH2")                   #Asparagine (Asn)
P <- Pro <- c("CH2","CH2","CH2")                  #Proline (Pro)
Q <- Gln <- c("CH2","CH2","CO","NH2")             #Glutamine (Gln)
R <- Arg <- c(rep("CH2",3), "NH", "CN+H2","NH2")  #Arginine (Arg)
S <- Ser <- c("CH2","OH")                         #Serine (Ser)
T <- Thr <- c("CH","OH","CH3")                   #Threonine (Thr)
V <- Val <- c("Methyl_CH","CH3")                  #Valine (Val)
W <- Trp <- c("CH2",Indole)                       #Tryptophan (Trp)
Y <- Tyr <- c("CH2", Phenol)                      #Tyrosine (Tyr)
```


Making a graphical representation of the amino acids.  Visited www.webmo.net/demoserver/cgi-bin/webmo/build_molecule.cgi using Mozilla FireFox v23.0 internet browser. Built the respective molecules from the list of fragments to build (Build -> Fragment... -> Category -> Amino Acid). If the Build option is grayed, the Build mode can be toggled using upside-down V icon on the left handside.  Orientated the Amino Acids with the carboxylic group (COO) in the back.  Then extracted the Cartesian coordinates from the Cartesian editor (Tools -> Cartesian Edit...).

```{r}

#Convert from subgroups to individual atoms
Mol2Atm = function(AA){
  AAcid <- gsub("_","",AA); AAcid <- gsub("Methyl","CH3",AAcid)
  AAcid <- gsub("2","H",AAcid); AAcid <- gsub("3","HH",AAcid)
  AAcid <- gsub("N+","X",AAcid,fixed=TRUE)
  AAcid <- gsub("O-","Z",AAcid,fixed=TRUE)
  AAcid <- strsplit(AAcid, split="")
  AAcid <- unlist(AAcid)
  AAcid <- gsub("X","N+",AAcid); AAcid <- gsub("Z","O-",AAcid)
  return(AAcid)
}

#Define path for conecting bonds in the Amino Acid
fpath <- function(AA){
  
  if(length(P) == length(AA) && all(P %in% AA) == TRUE
     && all(AA %in% P) == TRUE){
    
    pAA <- c(c(1,2,2,3,3,2,2,4,8),c(1,1,5,5,6),
             c(1,9,9,12,12,15,15,5,5,7),c(10,9,9,11),
             c(13,12,12,14),c(17,15,15,16))
  }
  
  else{
    if(length(A) == length(AA) && all(A %in% AA) == TRUE){
      pAA <- c(c(1,10,10,11),c(12,10,10,13))
    }
  
    if(length(C) == length(AA) && all(C %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14),c(11,10,10,12))
    }
  
    if(length(D) == length(AA) && all(D %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14,14,13,13,15),
               c(11,10,10,12))
    }
  
    if(length(E) == length(AA) && all(E %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,16,16,17,17,16,16,18),
               c(11,10,10,12),c(14,13,13,15))
    }
  
    if(length(F) == length(AA) && all(F %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14,14,16,16,18,18,20,20,22,22,13,13,22,22,23),
               c(11,10,10,12),c(15,14,14,16,16,17),c(19,18,18,20,20,21))
    }
    
    if(length(G) == length(AA) && all(G %in% AA) == TRUE){
      pAA <- c(1,10)
    }
  
    if(length(H) == length(AA) && all(H %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14,14,16,16,18,18,20,20,13,13,14,14,15),
               c(11,10,10,12),c(16,17),c(19,18,18,20,20,21))
    }
  
    if(length(I) == length(AA) && all(I %in% AA) == TRUE){
      pAA <- c(c(1,14,14,15),c(11,10,10,14,14,16,16,19,19,20),
               c(12,10,10,13),c(17,16,16,18),c(21,19,19,22))
    }
  
    if(length(K) == length(AA) && all(K %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,16,16,19,19,22,22,25),c(11,10,10,12), 
             c(14,13,13,15),c(20,19,19,21),c(23,22,22,24))
    }
  
    if(length(L) == length(AA) && all(L %in% AA) == TRUE){
      pAA <- c(c(1,10,10,17,17,13,13,14),c(11,10,10,12),c(15,13,13,16),
               c(18,17,17,19,19,21),c(20,19,19,22))
    }
  
    if(length(M) == length(AA) && all(M %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,16,16,17,17,18),c(11,10,10,12), 
               c(14,13,13,15),c(20,17,17,19))
    }
  
    if(length(N) == length(AA) && all(N %in% AA) == TRUE){
     pAA <- c(c(1,10,10,13,13,14,14,13,13,15,15,16),
              c(11,10,10,12),c(17,15))
    }  
  
    if(length(Q) == length(AA) && all(Q %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,16,16,17,17,16,16,18,18,20),
               c(11,10,10,12),c(14,13,13,15),c(19,18))
    }
  
    if(length(R) == length(AA) && all(R %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,16,16,19,19,20),c(23,22,22,21,21,19), 
               c(24,22,22,21,21,25),c(26,25,25,27),c(11,10,10,12),
               c(14,13,13,15),c(17,16,16,18))
    }
  
    if(length(S) == length(AA) && all(S %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14),c(11,10,10,12))
    }

    if(length(T) == length(AA) && all(T %in% AA) == TRUE){
      pAA <- c(c(1,10,10,14,14,15),c(11,10,10,12,12,13),
               c(16,14,14,17))
    }
  
    if(length(V) == length(AA) && all(V %in% AA) == TRUE){
      pAA <- c(c(1,14,14,10,10,13),c(11,10,10,12),
               c(15,14,14,16,16,19),c(17,16,16,18))
    }
  
    if(length(W) == length(AA) && all(W %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14,14,13,13,19,19,20),c(20,22,22,24),
               c(24,26,26,18,18,19),c(11,10,10,12),c(19,18,18,16,16,17),
               c(15,14,14,16),c(21,20,20,22,22,23),c(25,24,24,26,26,27))
    }
   
    if(length(Y) == length(AA) && all(Y %in% AA) == TRUE){
      pAA <- c(c(1,10,10,13,13,14,14,16,16,18,18,21),c(21,23,23,13),
               c(13,14,14,15),c(17,16,16,18,18,19,19,20),
               c(22,21,21,23,23,24))
    }
    AABbPath <- c(1,2,2,3,3,2,2,4,9,1,1,5,5,6,7,5,5,8)
    AAPath <- c(AABbPath,pAA)
    return(AAPath)
  }
}

#Define Vertices and bonds for Amino Acid Backbone (all but the 'R' sidechain)

VizAA = function(AA){                             ##visualize the amino acid
  AABackbone <- c("C","C","O","O-","N+",rep("H",4))
  

  if(length(P) == length(AA) && all(P %in% AA) == TRUE  
     && all(AA %in% P) == TRUE){
    
    gAA <- graph.empty(directed=FALSE) + vertices(AABackbone[1:8],Mol2Atm(AA))
    gAA <- gAA + edges(fpath(P))
  }
  else{
    gAA <- graph.empty(directed=FALSE) + vertices(AABackbone,Mol2Atm(AA))
    gAA <- gAA + edges(fpath(AA))
  }
}

#Generate a graphical representation of the Amino Acid

gAla <- graph.empty(directed=FALSE) + vertices("C","C",rep("H",3),"N+",rep("H",3),"C","O","O-","H")
gAla <- gAla + path(5,2,4) + path(3,2,1,6,9) + path(7,6,8) + path(13,1,10,11) + path(11,10,12)
layout <- matrix(c(0,0,0,1.5215,0,0,1.8875,1.0242,0,1.88710,-.51530,-0.88510,1.884,-.513,.8904,-.4972,.7077,1.2161,-.15040,1.6829,1.2197,-1.5319,.7112,1.2183,-.15510,.2212,2.0645,-.5073,-1.4149,-.002,-1.2091,-1.8122,-.9158,-.2057,-2.1995,.95010,-.35860,.51470,-.8867),ncol=3,byrow=TRUE)
gAla$layout <- layout.norm(layout)
rglplot(gAla)
```

```{r}
#Classifying AA into groups

AAclassifier <- function(AA){
  
  AAclass= data.frame(size,hydrophobic,polar,charge,aromatic)
  n = length(AA)
 
  if(n <= 3){                   #classifier for Small
                                #and Tiny AA residues
   AAclass$size <- "Small"     
    
   if(n < 3){
      
     AAclass$size <-"Tiny"
   }
  }
  else{
    AAclass$size <- "Not Small"
  }
  
  if(!is.na(match("C",AA)) == TRUE) {     #classifier for Aromaticity
    
    if(AA[match("C",AA):(length(Phenyl)+1)] == Phenyl ||
       AA[match("C",AA):(length(Imidazole)+1)] == Imidazole ||
       AA[match("C",AA):(length(Indole)+1)] == Indole ||
       AA[match("C",AA):(length(Phenol)+1)] == Phenol){
          
         AAclass$aromatic <- "Aromatic"
    }
  }
  else{
    AAclass$aromatic <- "Not Aromatic"
  }

  if(length(grep("OH|SH|NH2",AA)) != 0 || length(grep("+",AA,fixed=TRUE)) !=0 ||
     length(grep("-",AA,fixed=TRUE)) !=0 || length(grep("C",AA)) == 0){
    
    if(!is.na(match("Methyl_CH",AA)) == TRUE){  #classifierf for Aliphatic
    AAclass$hydrophobic <- c("Hydrophobic","Aliphatic")    
    }
  }
  if(length(grep("OH|+|-",AA,fixed=TRUE)) != 0){      #classifier for polar
    
    AAclass$polar = "Polar"
    
    if(!is.na(match("COO-",AA)) == TRUE){ #classifier for neg. charge
    
      AAclass$charge <- "Charge(-)"
    }
  
    if(length(grep("+",AA,fixed=TRUE)) != 0){ #classifier for pos. charge
    
      AAclass$charge <- "Charge(+)"
    }
  }
  return(AAclass)
}
```

```{r}
library(selectr)
#Acquire Mutant Frequency Table
 html("http://www.abren.net/protherm/mutation.html")%>%
+ html_node("table")%>%
+ html_table(fill=T) -> aATable

#---------------------------------------------------------------------------

#Acquire Further Data
url <- "http://www.abren.net/protherm/protherm_result.php?&DISP01=1&DISP02=1&DISP10=1&DISP14=1&DISP16=1&DISP18=1&DISP19=1&DISP22=1&DISP23=1&DISP24=1&DISP26=1&DISP30=1&DISP31=1&DISP38=1&NUMBER2=3000000&SOURCE=human&MUTA-C0=on&ASA-C=0-Any&ASA3=0-%&PH1=6.8&PH2=8.2&COUNT_E2=980&COUNT_S2=0"

xp <- '/html/body/table[2]'

AASearch <- url %>%
  html() %>%
  html_nodes(xpath=xp) %>%
  html_table(fill=TRUE)
AASearch <- AASearch[[1]]
AASearch <- AASearch[-c(16:21)]
names(AASearch) <- c("Entry","Protein","Source","Mutation","dG_H2O","dG","T", "Tm","dHcal","m","Cm","pH","Measure","Method",	"REFERENCE")

#--------------------------------------------------------------------------

url<- "http://p53.iarc.fr/AAProperties.aspx"
  
xp <- '//*[@id="aapropform"]/table'

AAProp <- url %>%
  html() %>%
  html_nodes(xpath=xp) %>%
  html_table(fill=TRUE)
AAProp <- AAProp[[1]]
AAProp <- AAProp[1:(nrow(AAProp)-1),]

#---------------------------------------------------------------------------

url <- "http://www.mhhe.com/physsci/chemistry/carey5e/Ch27/ch27-1-4-2.html"

AAPI <- url %>%
  html() %>%
  html_nodes("table") %>%
  html_table(fill=TRUE)                  #generates the table from the webpage
AAPI <- AAPI[[2]]

names_AAPI <- AAPI[1,]                   #extracts the names from the first row
names_AAPI <- gsub("Â|","",names_AAPI)    #removes Â characters (Å = alt 0194)
names_AAPI <- gsub("[[:space:]]","",names_AAPI)  #remove all the spaces
names_AAPI <- sub("oa","o a",names_AAPI)

AAPI <- AAPI[2:21,]                   #extracts the amino acid data   
AAPI <- gsub("Â| ","",AAPI)             
AAPI <- gsub("\"","",AAPI)            #removes all the \'s
AAPI <- sub("c","",AAPI)              #removes the 1st c from each line
AAPI <- gsub("(","",AAPI,fixed=TRUE)  #removes all the ('s
AAPI <- gsub(")","",AAPI,fixed=TRUE)  #removes all the )'s
AAPI <- strsplit(AAPI,split=",")      #splits the character arrays
AAPI <- as.data.frame(AAPI)           #remakes the data frame/table
names(AAPI) <- names_AAPI             #replaces the name with the original
```

```{r}
sessionInfo()
```